---
title: "Introduction to PxWebApiData"
author: "Øyvind Langsrud, Jan Bruusgaard, Solveig Bjørkholt and Susie Jentoft"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to PxWebApiData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
keywords: Statbank, PxWeb, PxWebApi, official statistics, json-stat
---


### Preface

An introduction to the R-package PxWebApiData is given below.
Six calls to the main function, `ApiData`, are demonstrated.
First, two calls for reading data sets are shown. The third call captures meta data.
However, in practise, one may look at the meta data first.
Then three more examples and some background is given.



```{r include = FALSE}
library(knitr)
library(PxWebApiData)
options(max.print = 36)

# Re-define the comment function to control line width and minimize excessive line breaks when printing.
comment <- function(x) {
     com <- base::comment(x)
     nchar_name <- min(103, 2 + max(nchar(com)))
     for (name in names(com)) {
         cat(strrep(" ", max(0, (nchar_name - nchar(name)))),
             name, 
             "\n", 
             strrep(" ", max(0, (nchar_name - nchar(com[[name]]) - 2))),
             "\"",
             com[[name]],
             "\"",  "\n", sep = "")
     }
 }
```


## Specification by variable indices and  variable id's

The dataset below has three variables, Region, ContentsCode and Tid. The variables can be used as input parameters.
Here two of the parameters are specified by variable id's
and one parameter is specified by indices. Negative values are used to specify reversed indices. Thus, we here obtain the two first and the two last years in the data.

A list of two data frames is returned; the label version and the id version.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",
        Region = c("1103", "0301"), ContentsCode = "Bosatte", Tid = c(1, 2, -2, -1))

```
To return a single dataset with only labels use the function `ApiData1`. The function `Apidata2` returns only id's. To return a dataset with both labels and id's in one dataframe use `ApiData12`.


```{r include = FALSE}
options(max.print = 75)
```


```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData12("http://data.ssb.no/api/v0/en/table/04861",
        Region = c("1103", "0301"), ContentsCode = "Bosatte", Tid = c(1, 2, -2, -1))

```


```{r include = FALSE}
options(max.print = 45)
```

## Specification by TRUE, FALSE and imaginary values (e.g. 3i).

All possible values is obtained by TRUE and corresponds to filter `"all": "*"` in the api query. Elimination of a variable is obtained by FALSE. An imaginary value corresponds to filter `"top"` in the api query.


```{r eval=TRUE, tidy = FALSE, comment=NA}

x <- ApiData("http://data.ssb.no/api/v0/en/table/04861",
        Region = FALSE, ContentsCode = TRUE, Tid = 3i)

```
To show either label version or id version
```{r eval=TRUE, tidy = FALSE, comment=NA}

x[[1]]

x[[2]]
```

## Show additional information

`comment` list additional dataset information: Title, latest update and source.
```{r, comment=NA}

comment(x)

```

## Obtaining meta data

Meta information about the data set can be obtained by `returnMetaFrames = TRUE`.


```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",  returnMetaFrames = TRUE)

```

## Aggregations using filter *agg:*
PxWebApi offers two more filters for groupings, `agg:` and `vs:`. You can see these filters in the code "API Query for this table" when you have made a table in PxWeb.

`agg`: is used for readymade aggregation groupings. 

This example shows the use of aggregation in age groups and aggregated timeseries for the new Norwegian municipality structure from 2020. Also note the url where `/en` is replaced by `/no`. That returns labels in Norwegian instead of in English.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/no/table/07459",
        Region = list("agg:KommSummer", c("K-3101", "K-3103")),
        Tid = 4i,
        Alder = list("agg:TodeltGrupperingB", c("H17", "H18")),
        Kjonn = TRUE)


```
There are two limitations in the PxWebApi using these filters.

1) The name of the filter and the id's are not shown in metadata, only in the code "API Query for this table".
2) The filters `agg:` and `vs:` can only take single elements as input. Filter `"all":"*"` eg. TRUE, does not work with agg: and vs:.

The other filter `vs:`, specify the grouping value sets, which is a part of the value pool. As it is only possible to give single elements as input, it is easier to query the value pool. This means that `vs:` is redundant.

In this example Region is the value pool and Fylker (counties) is the value set. As `vs:Fylker` is redundant, both will return the same:

```{r eval=TRUE, tidy = FALSE, comment=NA}
  Region = list("vs:Fylker",c("01","02"))
  Region = list(c("01","02"))

```

## Return the API query as JSON
In PxWebApi the original query is formulated as JSON. Using the parameter returnApiQuery is useful for debugging.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",  returnApiQuery = TRUE)

```

To convert an original JSON API query to a PxWebApiData query there is also a simple webpage [PxWebApiData call creator](https://radbrt-px-converter-px-converter-3mdstf.streamlit.app).

## Readymade datasets by GetApiData
Statistics Norway also provides an API with readymade datasets, available by http GET.
The data is most easily retrieved with the `GetApiData` function, which is the same as using  the parameter  `getDataByGET = TRUE` in the `ApiData` function.
This dataset is from Statistics Norway's Economic trends forecasts.

```{r eval=TRUE, comment=NA, tidy=FALSE}
x <- GetApiData("https://data.ssb.no/api/v0/dataset/934516.json?lang=en")
x[[1]]
comment(x)

```

## Eurostat data
Eurostat REST API offers JSON-stat version 2. It is possible to use this package to obtain data from Eurostat by using `GetApiData`
or the similar functions with `1`, `2` or `12` at the end

This example shows HICP total index, latest two periods for EU and Norway. See [Eurostat](https://wikis.ec.europa.eu/display/EUROSTATHELP/API+Statistics+-+data+query) for more.

```{r eval=TRUE, tidy = FALSE, comment=NA, encoding = "UTF-8"}

urlEurostat <- paste0(   # Here the long url is split into several lines using paste0 
  "https://ec.europa.eu/eurostat/api/dissemination/statistics/1.0/data/prc_hicp_mv12r", 
  "?format=JSON&lang=EN&lastTimePeriod=2&coicop=CP00&geo=NO&geo=EU")
urlEurostat
GetApiData12(urlEurostat)

```


## Practical example

We would like to extract the number of female R&D personel in the services sector of the Norwegian business life for the years 2019 and 2020.

1) Locate the relevant table at https://www.ssb.no that contains information on R&D personel. Having obtained the relevant table, table 07964, we create the link  https://data.ssb.no/api/v0/no/table/07964/


2) Load the package.

```{r, comment=NA, encoding = "UTF-8"}

library(PxWebApiData)

```


3) Check which variables that exist in the data.

```{r, comment=NA, encoding = "UTF-8"}

variables <- ApiData("https://data.ssb.no/api/v0/no/table/07964/",
                     returnMetaFrames = TRUE)

names(variables)

```

4) Check which values each variable contains.

```{r, comment=NA, encoding = "UTF-8"}

values <- ApiData("https://data.ssb.no/api/v0/no/table/07964/",
                  returnMetaData = TRUE)

values[[1]]$values
values[[2]]$values
values[[3]]$values

```

5) Define these variables in the query to sort out the values we want.

```{r, comment=NA, encoding = "UTF-8"}

mydata <- ApiData("https://data.ssb.no/api/v0/en/table/07964/",
                Tid = c("2021", "2022"), # Define year to 2021 and 2022
                NACE2007 = "G-N", # Define the services sector
                ContentsCode = c("KvinneligFoUpers")) # Define women R&D personell

mydata <- mydata[[1]] # Extract the first list element, which contains full variable names.

head(mydata)

```

6) Show additional information.

```{r, comment=NA, encoding = "UTF-8"}

comment(mydata)

```

## Background

PxWeb and it's API, PxWebApi is used as output database (Statbank) by many statistical agencies in the Nordic countries and several others, i.e. Statistics Norway, Statistics Finland, Statistics Sweden. See [list of installations](https://www.scb.se/en/services/statistical-programs-for-px-files/px-web/pxweb-examples/). 

For hints on using PxWebApi in general see [PxWebApi User Guide](https://www.ssb.no/en/api/pxwebapi/_/attachment/inline/3031ae43-a881-4ae6-b4c9-c04e190b1504:df8c31920354e37f30e21be5641df2d93a16ef6c/Api_user_manual.pdf).
