---
title: "Introduction to PxWebApiData"
author: "Øyvind Langsrud, Jan Bruusgaard, Solveig Bjørkholt and Susie Jentoft"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to PxWebApiData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
keywords: Statbank, PxWebApi, official statistics
---


### Preface

An introduction to the r-package PxWebApiData is given below.
Six calls to the main function, `ApiData`, are demonstrated.
First, two calls for reading data sets are shown. The third call captures meta data.
However, in practise, one may look at the meta data first.
Then three more examples and some background is given.



```{r include = FALSE}
library(knitr)
library(PxWebApiData)
options(max.print = 36)
```


## Specification by variable indices and  variable id's

The dataset below has three variables, Region, ContentsCode and Tid. The variables can be used as input parameters.
Here two of the parameters are specified by variable id's
and one parameter is specified by indices. Negative values are used to specify reversed indices. Thus, we here obtain the two first and the two last years in the data.

A list of two data frames is returned; the label version and the id version.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",
        Region = c("1103", "0301"), ContentsCode = "Bosatte", Tid = c(1, 2, -2, -1))

```
To return a single dataset with only labels use the function `ApiData1`. `Apidata2` returns only ids. To return a dataset with both labels and id's in one dataframe use `ApiData12`.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData12("http://data.ssb.no/api/v0/en/table/04861",
        Region = c("1103", "0301"), ContentsCode = "Bosatte", Tid = c(1, 2, -2, -1))

```


## Specification by TRUE, FALSE and imaginary values (e.g. 3i).

All possible values is obtained by TRUE and corresponds to filter `"all": "*"` in the api query. Elimination of a variables is obtained by FALSE. An imaginary value corresponds to filter `"top"` in the api query.


```{r eval=TRUE, tidy = FALSE, comment=NA}

x <- ApiData("http://data.ssb.no/api/v0/en/table/04861",
        Region = FALSE, ContentsCode = TRUE, Tid = 3i)

```
To show either label version or id version
```{r eval=TRUE, tidy = FALSE, comment=NA}

x[[1]]

x[[2]]
```


## Obtaining meta data

Meta information about the data set can be obtained by "returnMetaFrames = TRUE".


```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",  returnMetaFrames = TRUE)

```

## Aggregations using filter *agg:*
PxWebApi has two more filters for groupings, `agg:` and `vs:`. You can see these filters in the code "API Query for this table" when you have made a table in PxWeb.

`agg`: is used for readymade aggregation groupings. 

This example shows the use of aggregation in age groups and aggregated timeseries for the new Norwegian municipality structure from 2020. Note the url where /en is replaced by /no. This return Norwegian labels instead of English.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/no/table/07459",
        Region = list("agg:KommSummer", c("K-3001", "K-3002")),
        Tid = 4i,
        Alder = list("agg:TodeltGrupperingB", c("H17", "H18")),
        Kjonn = TRUE)


```
There are two limitations in the PxWebApi here.

1) The name of the filter and the IDs are not shown in metadata, only in the code "API Query for this table".
2) It is only possible to give single elements as input. Filter `"all":"*"` eg. TRUE, does not work with agg: and vs:.

The other filter `vs:`, specify the grouping value sets, which is a part of the value pool. As it is only possible to give single elements as input, it is easier to query the value pool. This means that vs: is redundant.

In this example Region is the value pool and Fylker is the value set. As `vs:Fylker` is redundant, both return the same:

```
  Region = list("vs:Fylker",c("01","02"))
  Region = list(c("01","02")).

```



## Return the API query as JSON
In PxWebApi the original query is formulated as JSON. Using the parameter returnApiQuery is useful for debugging.

```{r eval=TRUE, tidy = FALSE, comment=NA}
ApiData("http://data.ssb.no/api/v0/en/table/04861",  returnApiQuery = TRUE)

```


## Readymade datasets
Statistics Norway also provide an API with readymade datasets, available by http GET.
Use the parameter getDataByGET = TRUE. By changing to `lang=no` int the URL you get the label version in Norwegian.

This dataset is from Economic trends forecasts.

```{r eval=TRUE, comment=NA, tidy=FALSE}
x <- ApiData("https://data.ssb.no/api/v0/dataset/934516.json?lang=en", getDataByGET = TRUE)
x[[1]]

```


## Practical example

We would like to extract the number of female R&D personel in the services sector of the Norwegian business life for the years 2017 and 2018.

1) Locate the relevant table at https://www.ssb.no that contains information on R&D personel. Having obtained the relevant table, table 07964, we create the link  https://data.ssb.no/api/v0/no/table/07964/


2) Load the package.

```{r}

library(PxWebApiData)

```


3) Check which variables that exist in the data.

```{r}

variables <- ApiData("https://data.ssb.no/api/v0/no/table/07964/",
                     returnMetaFrames = TRUE)

names(variables)

```

4) Check which values each variable contains.

```{r}

values <- ApiData("https://data.ssb.no/api/v0/no/table/07964/",
                  returnMetaData = TRUE)

values[[1]]$values
values[[2]]$values
values[[3]]$values

```

5) Define these variables in the query to sort out the values we want.

```{r}

data <- ApiData("https://data.ssb.no/api/v0/en/table/07964/",
                Tid = c("2017", "2018"), # Define year to 2017 and 2018
                NACE2007 = "G-N", # Define the services sector
                ContentsCode = c("KvinneligFoUpers")) # Define women R&D personell

data <- data[[1]] # Extract the first list element, which contains full variable names.

head(data)

```



## Background

PxWeb and it's API, PxWebApi is used as output database (Statbank) by many statistical agencies in the Nordic countries and several others, i.e. Statistics Norway, Statistics Finland, Statistics Sweden. See [list of installations](https://www.scb.se/en/services/statistical-programs-for-px-files/px-web/pxweb-examples/). 

For hints on using PxWebApi in general see [PxWebApi User Guide](https://www.ssb.no/en/api/pxwebapi/_/attachment/inline/3031ae43-a881-4ae6-b4c9-c04e190b1504:df8c31920354e37f30e21be5641df2d93a16ef6c/Api_user_manual.pdf).
